A cascaded supervised learning approach to inverse reinforcement learning

### 2013-06-26 piot2012cascading_ECML/Cascading.tex
Modifications après revues ECML

\begin{document}
\title{A cascaded supervised learning approach to inverse reinforcement learning}
\titlerunning{Cascading approach to IRL}
\authorrunning{Klein, Piot, Geist and Pietquin}
\author{Edouard Klein\inst{1,}\inst{2}, Bilal Piot\inst{2,}\inst{3}, Matthieu Geist\inst{2}, Olivier Pietquin\inst{2,}\inst{3}}

\institute{
 1. ABC Team\\
 LORIA-CNRS, France.\\
\and
2. Supélec\\
 IMS-MaLIS Research group, France\\
 \texttt{firstname.lastname@supelec.fr}\\
\and 
3. UMI 2958 (GeorgiaTech-CNRS)\\
France\\}

\date{\today}

\maketitle

\begin{abstract}
  This paper considers the Inverse Reinforcement Learning (IRL) problem, that is inferring a reward function for which a demonstrated expert policy is optimal.
We propose to break the IRL problem down into two generic Supervised Learning steps: this is the Cascaded Supervised IRL (CSI) approach. A classification step that defines a score function is followed by a regression step providing a reward function.
A theoretical analysis shows that the demonstrated expert policy is near-optimal for the computed reward function.
Not needing to repeatedly solve a Markov Decision Process (MDP) and the ability to leverage existing techniques for classification and regression are two important advantages of the CSI approach. It is furthermore empirically demonstrated to compare positively to state-of-the-art approaches when using only transitions sampled according to the expert policy, up to the use of some heuristics. This is exemplified on two classical benchmarks (the mountain car problem and a highway driving simulator).
  \end{abstract}
\section{Introduction}
\label{sec-2}
Sequential decision making consists in choosing the appropriate action given the available data in order to maximize a certain criterion. When framed in a Markov Decision Process (MDP) (see Sec.~\ref{sec:background}), (Approximate) Dynamic programming ((A)DP) or Reinforcement Learning (RL) are often used to solve the problem by maximizing the expected sum of discounted rewards. The Inverse Reinforcement Learning (IRL)~\cite{russell1998learning} problem, which is addressed here, aims at inferring a reward function for which a demonstrated expert policy is optimal.

@@

It is possible to get imaginative in the last step. For example, using a Gaussian process regressor \cite{rasmussen2006gaussian} that outputs both expectation and variance can enable (notwithstanding a nontrivial amount of work) the use of reward-uncertain reinforcement learning \cite{regan2011robust}.

Our complete instantiation of CSI is summed up in Alg.~\ref{alg:CSIimp}.
\begin{algorithm}%[H]
    %\small
  \caption{A CSI instantiation with heuristics}
  \label{alg:CSIimp}
  \emph{\textbf{Given}} a dataset  $D_E = (s_k,a_k=\pi_E(a_k),s'_k)_k$\;\\
  \emph{\textbf{Construct}} the dataset $D_C=\{(s_i=s_k,a_i=\pi^E(s_i))=a_k\}$ \;\\
  \emph{\textbf{Train}} a score function-based classifier on $D_C$, obtaining decision rule $\pi_C$ and score function $q:S\times A \rightarrow \mathbb R$ \;\\
\emph{\textbf{Construct}} the dataset $\{((s_{j}=s_k,a_{j}=a_k),\hat r_j)_j\}$ with $\hat r_j = q(s_j,a_j) - \gamma q(s'_j=s'_k,\pi_C(s'_j=s'_k))$\;\\
\emph{\textbf{Set}} $r_{min} = \min_j\hat r_j - 1.$\;\\
\emph{\textbf{Construct}} the training set  $D_R = \{((s_{j}=s_k,a_{j}=a_k),\hat r_j)_j\}\cup\{((s_j=s_k,a),r_{min})_{j;\forall a\neq \pi_E(s_j) = a_k}\}$ \;\\
\emph{\textbf{Learn}} a reward function $\hat R^C$ from the training set $D_R$\;\\
\emph{\textbf{Output}} the reward function $\hat R^{C}: (s,a) \mapsto \omega^T \phi(s,a)$ \;
\end{algorithm}
\subsection{Mountain car}
\label{subsec:mountaincar}
The mountain car is a classical toy problem in RL: an underpowered car is tasked with climbing a steep hill. In order to do so, it has to first move away from the target and climb the slope on its left, and then it moves right, gaining enough momentum to climb the hill on the right on top of which lies the target. We used standard parameters for this problem, as can for example be found in \cite{sutton1998reinforcement}. When training an RL agent, the reward is, for example, $1$ if the car's position is greater than $0.5$ and $0$ anywhere else. The expert policy was a very simple hand crafted policy that uses the power of the car to go in the direction it already moves ({\it i.e.}, go left when the speed is negative, right when it is positive).\\

@@

The performance criterion for a policy $\pi$ is the mean (over the uniform distribution) value function with respect to $R_E$ : $\mathbf{E}_{s\sim\mathcal{U}}[V^{\pi}_{R_E}(s)]$.

Expert performance averages to $7.74$ ; we also show the natural random baseline that consists in drawing a random reward vector (with a uniform law) and training an agent on it. The reward functions found by SCIRL, CSI and RE are then optimized using a DP algorithm. The dataset $D_{random}$ needed by RE (and neither by CSI nor SCIRL) is made of $100$ episodes of length $10$ starting randomly in the state space and following a random policy. the dataset $D_E$ is made of $n$ episodes of length $n$, with $n \in \{3,7,10,15,20\}$.

@@

It is very interesting to compare our algorithm to the behavior of a classifier alone (respectively red and green plots on Fig.~\ref{tutu}). With \emph{the exact same data}, albeit the use of a very simple heuristics, the cascading approach demonstrates far better performance from the start. This is a clear illustration of the fact that using the Bellman equation to construct the data fed to the regressor and outputting not a policy, but a reward function that can be optimized on the MDP truly makes use of the information that the transitions $(s,a,s')$ bear (we recall that the classifier only uses $(s,a)$ couples).

Furthermore, the classifier whose results are displayed here is the output of the first step of the algorithm. The classification performance is obviously not that good, which points to the fact that our algorithm may be empirically more forgiving of classification errors than our theoretical bound lets us expect.

>>>>

\setlength\floatsep{1.25\baselineskip plus 3pt minus 2pt}
\setlength\textfloatsep{1.25\baselineskip plus 3pt minus 2pt}
\setlength\intextsep{1\baselineskip plus 3pt minus 2 pt}

\begin{document}
\title{A cascaded supervised learning approach to inverse reinforcement learning}
\titlerunning{Cascading approach to IRL}
\authorrunning{Klein, Piot, Geist and Pietquin}
\author{Edouard Klein\inst{1,}\inst{2}, Bilal Piot\inst{2,}\inst{3}, Matthieu Geist\inst{2}, Olivier Pietquin\inst{2,}\inst{3}\thanks{The research leading to these results has received funding from the European Union Seventh Framework Programme (FP7/2007-2013) under grant agreement n\textdegree270780.}}

\institute{
 ABC Team\, LORIA-CNRS, France.\\
\and
Supélec, IMS-MaLIS Research group, France\\
 \texttt{firstname.lastname@supelec.fr}\\
\and
UMI 2958 (GeorgiaTech-CNRS), France\\
}

%\date{\today}

\maketitle
%\renewcommand*\abstractname{Acknowledgement.}
%\begin{abstract}
%The research leading to these results has received funding from the European Union Seventh Framework Programme (FP7/2007-2013) under grant agreement n\textdegree270780.
%\end{abstract}

\renewcommand*\abstractname{Abstract.}
\begin{abstract}
  This paper considers the Inverse Reinforcement Learning (IRL) problem, that is inferring a reward function for which a demonstrated expert policy is optimal.
We propose to break the IRL problem down into two generic Supervised Learning steps: this is the Cascaded Supervised IRL (CSI) approach. A classification step that defines a score function is followed by a regression step providing a reward function.
A theoretical analysis shows that the demonstrated expert policy is near-optimal for the computed reward function.
Not needing to repeatedly solve a Markov Decision Process (MDP) and the ability to leverage existing techniques for classification and regression are two important advantages of the CSI approach. It is furthermore empirically demonstrated to compare positively to state-of-the-art approaches when using only transitions sampled according to the expert policy, up to the use of some heuristics. This is exemplified on two classical benchmarks (the mountain car problem and a highway driving simulator).
  \end{abstract}
\section{Introduction}
\label{sec-2}
Sequential decision making consists in choosing the appropriate action given the available data in order to maximize a certain criterion. When framed in a Markov Decision Process (MDP) (see Sec.~\ref{sec:background}), (Approximate) Dynamic programming ((A)DP) or Reinforcement Learning (RL) are often used to solve the problem by maximizing the expected sum of discounted rewards. The Inverse Reinforcement Learning (IRL)~\cite{russell1998learning} problem, which is addressed here, aims at inferring a reward function for which a demonstrated expert policy is optimal.

@@

It is possible to get imaginative in the last step. For example, using a Gaussian process regressor \cite{rasmussen2006gaussian} that outputs both expectation and variance can enable (notwithstanding a nontrivial amount of work) the use of reward-uncertain reinforcement learning \cite{regan2011robust}. Our complete instantiation of CSI is summed up in Alg.~\ref{alg:CSIimp}.
\begin{algorithm}%[H]
    %\small
  \caption{A CSI instantiation with heuristics}
  \label{alg:CSIimp}
  \emph{\textbf{Given}} a dataset  $D_E = (s_k,a_k=\pi_E(a_k),s'_k)_k$\;\\
  \emph{\textbf{Construct}} the dataset $D_C=\{(s_i=s_k,a_i=\pi^E(s_i))=a_k\}$ \;\\
  \emph{\textbf{Train}} a score function-based classifier on $D_C$, obtaining decision rule $\pi_C$ and score function $q:S\times A \rightarrow \mathbb R$ \;\\
\emph{\textbf{Construct}} the dataset $\{((s_{j}=s_k,a_{j}=a_k),\hat r_j)_j\}$ with $\hat r_j = q(s_j,a_j) - \gamma q(s'_j=s'_k,\pi_C(s'_j=s'_k))$\;\\
\emph{\textbf{Set}} $r_{min} = \min_j\hat r_j - 1.$\;\\
\emph{\textbf{Construct}} the training set  $D_R = \{((s_{j}=s_k,a_{j}=a_k),\hat r_j)_j\}\cup\{((s_j=s_k,a),r_{min})_{j;\forall a\neq \pi_E(s_j) = a_k}\}$ \;\\
\emph{\textbf{Learn}} a reward function $\hat R^C$ from the training set $D_R$\;\\
\emph{\textbf{Output}} the reward function $\hat R^{C}: (s,a) \mapsto \omega^T \phi(s,a)$ \;
\end{algorithm}
\subsection{Mountain car}
\label{subsec:mountaincar}
The mountain car is a classical toy problem in RL: an underpowered car is tasked with climbing a steep hill. In order to do so, it has to first move away from the target and climb the slope on its left, and then it moves right, gaining enough momentum to climb the hill on the right on top of which lies the target. We used standard parameters for this problem, as can for example be found in \cite{sutton1998reinforcement}. When training an RL agent, the reward is, for example, $1$ if the car's position is greater than $0.5$ and $0$ anywhere else. The expert policy was a very simple hand crafted policy that uses the power of the car to go in the direction it already moves ({\it i.e.}, go left when the speed is negative, right when it is positive).\\

@@

The performance criterion for a policy $\pi$ is the mean (over the uniform distribution) value function with respect to $R_E$ : $\mathbf{E}_{s\sim\mathcal{U}}[V^{\pi}_{R_E}(s)]$. Expert performance averages to $7.74$ ; we also show the natural random baseline that consists in drawing a random reward vector (with a uniform law) and training an agent on it. The reward functions found by SCIRL, CSI and RE are then optimized using a DP algorithm. The dataset $D_{random}$ needed by RE (and neither by CSI nor SCIRL) is made of $100$ episodes of length $10$ starting randomly in the state space and following a random policy. the dataset $D_E$ is made of $n$ episodes of length $n$, with $n \in \{3,7,10,15,20\}$.

@@

It is very interesting to compare our algorithm to the behavior of a classifier alone (respectively red and green plots on Fig.~\ref{tutu}). With \emph{the exact same data}, albeit the use of a very simple heuristics, the cascading approach demonstrates far better performance from the start. This is a clear illustration of the fact that using the Bellman equation to construct the data fed to the regressor and outputting not a policy, but a reward function that can be optimized on the MDP truly makes use of the information that the transitions $(s,a,s')$ bear (we recall that the classifier only uses $(s,a)$ couples). Furthermore, the classifier whose results are displayed here is the output of the first step of the algorithm. The classification performance is obviously not that good, which points to the fact that our algorithm may be empirically more forgiving of classification errors than our theoretical bound lets us expect.
